---
title: "Group 2"
author: "Ben Aland, Birch Lazo-Murphy, Darcy Perin, Nick Steichmann"
date: "4/14/2022"
output: html_document
---

Individual codes, made figures, description and rationale for analysis.

### Ben's Code
```{r echo = FALSE, warning=FALSE, message=FALSE}

```

### Birch's Code
```{r echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(viridisLite)
library(viridis)

data18 <- read_xls("data/2018303d_final.xls")
data08 <- read_xls("data/SC_303d_lists_2006to2016/2008303dfinal020608prtyrk.xls")
data98 <- read_xls("data/SC_303d_lists_1998to2004/1998303dfin_al_rec_only.xls")

# Stations who FAILED to improve from 1998 to 2008
stationsFail98to08 <- inner_join(data98, data08, by = "STATION")
Failed98to08MoreThanOne <- stationsFail98to08 %>% 
  group_by(LOCATION) %>% 
  summarize(count=n())

# Stations who FAILED to improve from 1998 to 2018
stationsFail20yr <- inner_join(stationsFail98to08, data18, by = "STATION")

# Stations who got better and were REMOVED from 1998 to 2018
stationsPass20yr <- data18[!data18$STATION %in% stationsFail20yr$STATION,]

# Stations who were ADDED between 1998 and 2008 
stationsWorse98to08 <- data08[!data08$STATION %in% data98$STATION,]

# Stations who were ADDED between 1998 and 2018
stationsWorse20yr <- data18[!data18$STATION %in% stationsWorse98to08$STATION,]

# Number of Causes for Each Location that has continuously failed
Failed98to18MoreThanOne <- stationsFail20yr %>% 
  group_by(LOCATION) %>% 
  summarize(NumCauses=n()) %>%
  group_by(NumCauses) %>%
  summarize(count2 = n())
Failed98to18MoreThanOne$NumCauses <- sub("^", "N", Failed98to18MoreThanOne$NumCauses)
Failed98to18MoreThanOne$Percent <- c(Failed98to18MoreThanOne$count2[1]/sum(Failed98to18MoreThanOne$count2)*100,
                                     Failed98to18MoreThanOne$count2[2]/sum(Failed98to18MoreThanOne$count2)*100,
                                     Failed98to18MoreThanOne$count2[3]/sum(Failed98to18MoreThanOne$count2)*100,
                                     Failed98to18MoreThanOne$count2[4]/sum(Failed98to18MoreThanOne$count2)*100,
                                     Failed98to18MoreThanOne$count2[5]/sum(Failed98to18MoreThanOne$count2)*100,
                                     Failed98to18MoreThanOne$count2[6]/sum(Failed98to18MoreThanOne$count2)*100)

ggplot(Failed98to18MoreThanOne, aes(x= "", y=Percent, fill=NumCauses)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_viridis(option = "C", discrete = TRUE) +
  theme_void() +
  labs(title = "Number of Causes for Locations Listed from 1998 to 2018")

# Number of Causes for Each Location that has been removed from the list
Pass98to18MoreThanOne <- stationsPass20yr %>% 
  group_by(DESCRIPTION) %>% 
  summarize(NumCauses=n()) %>%
  na.omit() %>%
  group_by(NumCauses) %>%
  summarize(count2 = n()) 
Pass98to18MoreThanOne$NumCauses <- sub("^", "N", Pass98to18MoreThanOne$NumCauses)
Pass98to18MoreThanOne$Percent <- c(Pass98to18MoreThanOne$count2[1]/sum(Pass98to18MoreThanOne$count2)*100,
                                     Pass98to18MoreThanOne$count2[2]/sum(Pass98to18MoreThanOne$count2)*100,
                                     Pass98to18MoreThanOne$count2[3]/sum(Pass98to18MoreThanOne$count2)*100)
                                     

ggplot(Pass98to18MoreThanOne, aes(x= "", y=Percent, fill=NumCauses)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_viridis(option = "C", discrete = TRUE) +
  theme_void() +
  labs(title = "Number of Causes for Locations Removed 1998 to 2018")
# They are doing a shit job of removing stations with more than one issue

# Number of Causes for Each Location that has been added
Add98to18MoreThanOne <- stationsWorse20yr %>% 
  group_by(DESCRIPTION) %>% 
  summarize(NumCauses=n()) %>%
  na.omit() %>%
  group_by(NumCauses) %>%
  summarize(count2 = n()) 
Add98to18MoreThanOne$NumCauses <- sub("^", "N", Add98to18MoreThanOne$NumCauses)
Add98to18MoreThanOne$Percent <- c(Add98to18MoreThanOne$count2[1]/sum(Add98to18MoreThanOne$count2)*100,
                                   Add98to18MoreThanOne$count2[2]/sum(Add98to18MoreThanOne$count2)*100)


ggplot(Add98to18MoreThanOne, aes(x= "", y=Percent, fill=NumCauses)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_viridis(option = "C", discrete = TRUE) +
  theme_void() +
  labs(title = "Number of Causes for Locations Added 1998 to 2018")

# Stations with a lot of failures
failures <- stationsFail20yr %>% 
  group_by(LOCATION) %>% 
  summarize(NumCauses=n()) %>%
  filter(NumCauses == 8)
```

The goal of these figures is to be able to easily visualize improvements to water bodies over time. The first pie chart shows the fraction of locations which have not been removed over the 20-year period 1998 to 2018. They are grouped by the number of causes which have caused each respective location to be on the list. The majority of locations have made the list due to a single issue, and have remained there for 20 years, however, in four locations over eight issues have persisted. These locations are 1) Magnolia Gardens, a former plantation, 2) Kingston Lake, which is next to a recycling plant, 3) McApline Creek, which is next to a construction facility that specializes in retention walls, 4) Reedy River, which is to the southeast of a wetland (nature's kidneys).The second pie chart shows the fraction of locations which have been removed over the 20-year period. The overwhelming majority of the water bodies have only had one issue to deal with. Water bodies with multiple issues are harder to remediate as they require more complex strategies. The final pie chart shows the fraction of water bodies that have been added over the 20-year period. Most of these water bodies only have one issue - this is likely just a function of increased monitoring of water bodies.

### Darcy's Code
```{r echo = FALSE, warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
```
```{r}
data_1998 = read_csv("data/dp_data/1998303dfin_al_rec_only_dp.csv", col_names = c("site", "station", "use", "cause", "HUC", "basin", "priority"))

data_1998_edit= data_1998[-c(1),]
```

```{r}
data_2008 = read_csv("data/dp_data/2008303dfinal020608_dp.csv", col_names = c("target_date", "note", "basin", "HUC", "location", "station", "county", "use", "cause"))

data_2008_edit= data_2008[-c(1),]
```

```{r}
data_2018 = read_csv("data/dp_data/2018303d_final_dp.csv", col_names = c("priority_rank", "note", "basin", "HUC", "county","description", "station", "use", "cause"))

data_2018_edit= data_2018[-c(1),]
```

```{r}
causes_98= data_1998_edit %>%
  filter(!is.na(cause))

causes_08= data_2008_edit %>%
  filter(!is.na(cause)) 

causes_18= data_2018_edit %>%
  filter(!is.na(cause)) 
```

```{r}
plot3=causes_98 %>%
  ggplot(aes(x = cause, fill=basin)) +
  geom_bar() +
  facet_grid(vars(basin)) +
  labs(x = "Pollutants", y = "Basins")+
  ggtitle("1998 Pollutant Data")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(strip.text.y = element_blank()) +
  scale_fill_manual(values = c("firebrick2", "darkorange1", "goldenrod1", "seagreen3", "forestgreen", "royalblue2", "darkorchid", "plum1")) +
   theme(legend.title=element_blank())

plot3
```

```{r}
plot1=causes_08 %>%
  ggplot(aes(x = cause, fill=basin)) +
  geom_bar() +
  facet_grid(vars(basin)) +
  labs(x = "Pollutants", y = "Basins")+
  ggtitle("2008 Pollutant Data")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(strip.text.y = element_blank()) +
  scale_fill_manual(values = c("firebrick2", "darkorange1", "goldenrod1", "seagreen3", "forestgreen", "royalblue2", "darkorchid", "plum1")) +
   theme(legend.title=element_blank())

plot1
```

```{r}
plot2=causes_18 %>%
  ggplot(aes(x = cause, fill=basin)) +
  geom_bar() +
  facet_grid(vars(basin)) +
  labs(x = "Pollutants", y = "Basins")+
  ggtitle("2018 Pollutant Data")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(strip.text.y = element_blank())+
  scale_fill_manual(values = c("firebrick2", "darkorange1", "goldenrod1", "seagreen3", "forestgreen", "royalblue2", "darkorchid", "plum1")) +
   theme(legend.title=element_blank()) 

  
plot2
```

```{r}
library(ggplot2)
library(ggpubr)
```

```{r}
ggarrange(plot3, plot1, plot2, ncol = 1, nrow = 3)
```

```{r}
NH3N_causes_08= causes_08 %>%
  filter(cause== "NH3N")%>%
  group_by(basin)
#8 Stations

NH3N_causes_18= causes_18 %>%
  filter(cause== "NH3N")%>%
  group_by(basin)
#5 Stations

NH3N_causes_98= causes_98 %>%
  filter(cause== "NH3N")%>%
  group_by(basin)
#0 Stations

CHLA_causes_08= causes_08 %>%
  filter(cause== "CHLA")
#15 Stations

CHLA_causes_18= causes_18 %>%
  filter(cause== "CHLA")
#25 Stations

CHLA_causes_98= causes_98 %>%
  filter(cause== "CHLA")
#0 Stations

CR_causes_08= causes_08 %>%
  filter(cause== "CR")
#3 Station

CR_causes_18= causes_18 %>%
  filter(cause== "CR")
#2 Station

CR_causes_98= causes_98 %>%
  filter(cause== "CR")
#9 Station

HG_causes_08= causes_08 %>%
  filter(cause== "MG")
#0 Stations

HG_causes_18= causes_18 %>%
  filter(cause== "MG")
#0 Stations

HG_causes_98= causes_98 %>%
  filter(cause== "MG")
#0 Stations

CU_causes_08= causes_08 %>%
  filter(cause== "CU")
#93 Stations

CU_causes_18= causes_18 %>%
  filter(cause== "CU")
#18 Stations

CU_causes_98= causes_98 %>%
  filter(cause== "CU")
#36 Stations

DO_causes_08= causes_08 %>%
  filter(cause== "DO")
#124 Stations

DO_causes_18= causes_18 %>%
  filter(cause== "DO")
#158 Stations

DO_causes_98= causes_98 %>%
  filter(cause== "DO")
#57 Stations

ENTERO_causes_08= causes_08 %>%
  filter(cause== "ENTERO")
#11 Stations

ENTERO_causes_18= causes_18 %>%
  filter(cause== "ENTERO")
#73 Stations

ENTERO_causes_98= causes_98 %>%
  filter(cause== "ENTERO")
#0 Stations

FC_causes_08= causes_08 %>%
  filter(cause== "FC")
#346 Stations

FC_causes_18= causes_18 %>%
  filter(cause== "FC")
#116 Stations

FC_causes_98= causes_98 %>%
  filter(cause== "FC")
#325 Stations

PH_causes_08= causes_08 %>%
  filter(cause== "PH")
#66 Stations

PH_causes_18= causes_18 %>%
  filter(cause== "PH")
#97 Stations

PH_causes_98= causes_98 %>%
  filter(cause== "PH")
#21 Stations

TN_causes_08= causes_08 %>%
  filter(cause== "TN")
#6 Stations

TN_causes_18= causes_18 %>%
  filter(cause== "TN")
#23 Stations

TN_causes_98= causes_98 %>%
  filter(cause== "TN")
#0 Stations

ECOLI_causes_08= causes_08 %>%
  filter(cause== "ECOLI")
#0 Stations

ECOLI_causes_18= causes_18 %>%
  filter(cause== "ECOLI")
#163 Stations

ECOLI_causes_98= causes_98 %>%
  filter(cause== "ECOLI")
#0 Stations

CD_causes_08= causes_08 %>%
  filter(cause== "CD")
#0 Stations

CD_causes_18= causes_18 %>%
  filter(cause== "CD")
#11 Stations

CD_causes_98= causes_98 %>%
  filter(cause== "CD")
#0 Stations

BIO_causes_08= causes_08 %>%
  filter(cause== "BIO")
#165 Stations

BIO_causes_18= causes_18 %>%
  filter(cause== "BIO")
#170 Stations

BIO_causes_98= causes_98 %>%
  filter(cause== "BIO")
#170 Stations

TURBIDITY_causes_08= causes_08 %>%
  filter(cause== "TURBIDITY")
#58 Stations

TURBIDITY_causes_18= causes_18 %>%
  filter(cause== "TURBIDITY")
#58 Stations

TURBIDITY_causes_98= causes_98 %>%
  filter(cause== "TURBIDITY")
#0 Stations

TP_causes_08= causes_08 %>%
  filter(cause== "TP")
#51 Stations

TP_causes_18= causes_18 %>%
  filter(cause== "TP")
#80 Stations

TP_causes_98= causes_98 %>%
  filter(cause== "TP")
#16 Stations

PB_causes_08= causes_08 %>%
  filter(cause== "PB")
#0 Stations

PB_causes_18= causes_18 %>%
  filter(cause== "PB")
#0 Stations

PB_causes_98= causes_98 %>%
  filter(cause== "PB")
#0 Stations

PCB_causes_08= causes_08 %>%
  filter(cause== "PCB")
#5 Stations

PCB_causes_18= causes_18 %>%
  filter(cause== "PCB")
#21 Stations

PCB_causes_98= causes_98 %>%
  filter(cause== "PCB")
#0 Stations

NI_causes_08= causes_08 %>%
  filter(cause== "NI")
#11 Stations

NI_causes_18= causes_18 %>%
  filter(cause== "NI")
#2 Stations

NI_causes_98= causes_98 %>%
  filter(cause== "NI")
#0 Stations

ZN_causes_08= causes_08 %>%
  filter(cause== "ZN")
#22 Stations

ZN_causes_18= causes_18 %>%
  filter(cause== "ZN")
#11 Stations

ZN_causes_98= causes_98 %>%
  filter(cause== "ZN")
#20 Stations

ORGANOTINS_causes_08= causes_08 %>%
  filter(cause== "ORGANOTINS")
#0 Stations

ORGANOTINS_causes_18= causes_18 %>%
  filter(cause== "ORGANOTINS")
#0 Stations

ORGANOTINS_causes_98= causes_98 %>%
  filter(cause== "ORGANOTINS")
#0 Stations

SED_causes_98= causes_98 %>%
  filter(cause== "SED")
#3 Stations

PEST_causes_98= causes_98 %>%
  filter(cause== "PEST")
#2 Stations
```

```{r}
library(dplyr)
library(kableExtra)

station= read_csv("data/dp_data/station_numb.csv")

station %>%
  kbl(caption = "Number of Stations with The Same Pollutants") %>%
  kable_classic(full_width = F, html_font = "Cambria") 
```

The objective of this code is to be able to visualize the major pollutants found in each basin for 1998, 2008 and 2018 using a ggplot. The second goal was to make a table stating the numerical values of the number of stations with the same pollutant for 1998, 2008, and 2018. This is important because it allows us to see if specific pollutants persist for the 20 year period. Based on the graphs, it can be seen that Fecal Coliform Bacteria is the greatest pollutant, but the amount of stations it effects has significantly decreased between 2008 and 2018. Another trend is that 2018 has the most variety of pollutants, and there is an increase in almost every pollutant between 2008 and 2018. This may be due to either and increase in pollution, or an increase in the amount of testing done to detect the pollutants. 


### Nick's Code
```{r echo = FALSE, warning=FALSE, message=FALSE}

```
